Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-06 WSSSPE3/report/WSSSPE3_review.tex
tweaks to KBase lightning talk description

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham, and
Sergei Maslov: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standard to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also didn't have standards for software
engineering. This is contrasted to computer science research, where these
approaches shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Instead of this more
traditional approach, KBase uses a variation on the ``Scrum'' methodology.
After picking projects and team members, four to five teams work on projects for about two
weeks before a one-week evaluation by the executive committee, and after which new
teams and projects might be chosen to restart the ``agile'' development cycle.
This process is intentionally open and documented.

>>>>

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham,
Sergei Maslov, and the KBase Project: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standards to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also does not have standards for software
engineering. This is contrasted to computer science research, where software
engineering standards shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Instead of this more
traditional approach, KBase uses a variation on the ``Scrum'' methodology.
After picking projects and team members, four to five teams work on projects for about two
weeks before a one-week evaluation by the executive committee. Based on their assessment new
teams and projects might be chosen before iteratively restarting the ``agile'' development cycle.
This process is intentionally open and documented.
